{\LARGE Basis Choice for Scalar-on-Function Regression \\ with Applications to Near-Infrared Spectroscopy}

### 2022-02-08 Paper/main.tex
Meeting changes

	\subsection{Results}	
	The discussed results and figures of $\hat{\beta}(t)$ for the simulation can be found in the appendix. 
	\subsubsection{Basis Expansion Regression}
	The following results origin from the \hyperref[basis_exp_transf]{Estimation using Basis-Representation}, in which we transform the observed functions to perform regression of a scalar on a countable sequence of regressors, which is then tractable with typical multivariate regression theory. examining the results, it appears that the cross-validated MSPE inherets a convex nature in the increasing basis functions.

	\paragraph{Monomial Basis}
	Due to the in the number of basis functions increasing collinearity of these basis functions, simulations were conducted up until the sixth monomial basis, which already shows signs of numerical instability. This could be solved using orthogonal polynomials like Legendre polynomials, see \cite{Dattoli_2001}. For reasons outlined in {\color{green} Link chapter}, they seem suited for $f_1$. The monomial basis shows a better performance than B-splines for this coefficient function. A hypothesis for this could be that $f_1$ resembles an entire function, which can be well approximated with a power series. In case of $f_2$, this basis shows the weakest performance out of all basis, which is visible in figure 9 and 10, where it seems like $\hat{\beta}$ is not changing in the amount of noise and shows exaggerated behaviour at the boundaries. For $f_1$, the simulation selects 5(3) and for $f_2$ 5(5) monomial basis functions for the high(low) signal-to-noise ratios. This weakness is especially pronounced in the setup $f_2,Y_1$; for the high noise setup, $f_2,Y_2$ this weakness is still visible, but less pronounced.
	\vspace{-0.2cm}

	\paragraph{B-spline Basis}
	Simulations with B-spline basis functions were possible from 5 to 18, since from 18 onwards the simulations were running into problems concerning {\color{green} what happened here exactly?}
	The smooth function $f_1$ requires only 5(4) B-spline basis functions, for the high(low) signal-to-noise ratio to obtain the best fit for the B-spline basis, which performs the worst for $f_1$. An explanation might be its exaggerated behaviour at the boundaries and the exaggeration of the peculiarities of $f_1$ (figure 7 and 8), which is especially pronounced for the higher noise responses. For $f_2$, 11(6) B-spline basis functions are chosen for the high(low) signal-to-noise ratios. For $f_2$, the B-spline basis functions outperform the monomial basis but come second to the Fourier basis. While the basis seems to recognize the peculiarities in $f_2,Y_1$, it seems to struggle with it for the noisy responses in $f_2,Y_2$ (figure 9 and 10).
	\vspace{-0.2cm}

	\paragraph{Fourier Basis}
	For $f_1$, the simulation chooses a smaller number of Fourier basis functions, 5(3) and a higher number for $f_2$, 9(7) for the setup with the high(low) signal-to-noise ratio. With the low signal-to-noise ratio, the simulation chooses a smaller number of basis functions, which could be to account for the higher noise in the response variable. The Fourier basis functions perform the best for each setup for the basis expansion regression. Several reasons could contribute to this: First, especially $f_1$ shows similar curvature order across the domain while the curvature of $f_2$, while changing, does not display any eradic jumps. Second, both functions feature periodic behaviour. Third, $f_2$ does have the same value at the start- and the end of the interval.	\\

	\subsubsection{Functional Principal Component Regression}
		The model which is used for the FPCR is described in \hyperref[fpc_exp_transf]{Estimation using Functional Principal Components}. Additionally to the selection of the truncation parameter $K$, the choice of the number of FPC's adds to the complexity of the model since the eigenvalues and -functions from the decomposition are influenced by the choice of $K$ that was used to approximate the function. But since the FPCR is ultimately estimated with a linear model with the FPC's as regressors, the degree of freedom of these models is not affected by $K$, but only by the number of FPC's $n_{FPC} \in \{2, 3, 4 \}$. As will be described, it seems that neither a convex behaviour of the MSPE, nor a clear relationship can be observed between the increasing number of basis function and the number of FPC's.

	\paragraph{Monomial Basis}
	For $f_1$, the cross-validated MSE is decreasing in the increasing number of FPC's and chosen basis functions for both the setup with the high, and the low signal-to-noise ratio (4,5,6 basis functions for $n_{FPC} \in \{2, 3, 4 \}$). In the setup using $f_2$ we also observe a decreasing MSE in the number of FPC, but a different behaviour of the chosen basis functions: While the model selects 10 basis functions for $f_2,Y_1$ and $f_2,Y_2$ using 2 FPC, it only selects 6 for the same two specifications using 3 FPC and 8 for $f_2,Y_1$ and $f_2,Y_2$ with 4 FPC. This might be surprising, since one might assume that the number of basis functions is increasing in $n_{FPC}$ since more basis functions led to less smoothing and therefore more variability in the expanded curves that might be exploited by the FPC's. The monomial basis shows the weakest performance for all three basis functions in each setting.
	\vspace{-0.2cm}

@@

	\paragraph{Fourier Basis}
	We observe similar behaviour as for the FPCR using B-splines. In $f_1,Y_1$, the MSE is decreasing in $n_{FPC}$ while the results in $f_1,Y_2$ might indicate overfitting when using four FPC's. Since it is not possible to calculate the cross-validated MSE for three basis functions for $n_{FPC}$ = 4, the increasing number of basis functions as seen in the FPCR with B-splines can not be observed here for sure. $f_2,Y_1$ displays the greatest relative decline of MSE in the increasing number of FPC's, but in absolut values, therefore acknowledging the higher noise responses for $f_2,Y2$, the decline if MSE is similar to the decline observed for $f_2,Y_2$. Both configurations of $f_2$ are using 5,15,7 basis functions for $n_{FPC} \in \{2, 3, 4 \}$: This behaviour in increasing $n_{FPC}$ opposes the previously reported behaviour of the monomial basis and the increasing nature of the B-splines basis functions.
	\vspace{-0.2cm}

@@

	\section{Application}\label{Application}
		The application uses the methods and insights from the previous sections to predict the octane ratings of the introduced gasoline dataset.
		Although the simulation study granted valuable insights in the different methods in four different settings, it is not enough to determine the method and choice of basis of the application, since there is too much uncertainty involved. To point out some sources of uncertainty: First, opposing to the simulation setup, we do not know the true coefficient function. Visual inspection fuels the hypothesis that $NIR$ looks closer to $f_2$ than to $f_1$ but the insights from two functions are not enough to draw any conclusion. Second, we have no information about the signal-to-noise ratios of the measured octane numbers. Third, we made assumptions about the distribution of $\xi$ that are not relateable to this real-life application where we not know this distribution, neither can we infer something about it. 
		Therefore, we will run all specification again for the gasoline dataset and use the results of the simulation study towards the reading of the results. The application is designed similar to the simulation study: The 60 spectra of the gasoline dataset will be used to regress them on the reported octane numbers and evaluate the results as prediction MSE using 12 fold cross-validation with 5 elements per fold. In total, we conducted 1000 repetitions for each setting.

>>>>

	\subsection{Results}	
	The discussed results and figures of $\hat{\beta}(t)$ for the simulation can be found in the appendix. 
	\subsubsection{Basis Expansion Regression}
	The following results origin from the \hyperref[basis_exp_transf]{Estimation using Basis-Representation}, in which we transform the smooth curves to perform regression of a scalar on a finite sequence of regressors, which is then tractable with typical multivariate regression theory. examining the results, it appears that the cross-validated MSPE inherets a convex nature in the increasing basis functions.

	\paragraph{Monomial Basis}
	Due to the in the number of basis functions increasing collinearity of these basis functions, simulations were conducted up until the sixth monomial basis, which already shows signs of numerical instability. This could be solved using orthogonal polynomials like Legendre polynomials, see \cite{Dattoli_2001}. For reasons outlined in {\color{green} Link chapter}, they seem suited for $f_1$. The monomial basis shows a better performance than B-splines for this coefficient function. A hypothesis for this could be that $f_1$ is an entire function, which can be well approximated with a power series. In case of $f_2$, this basis shows the weakest performance out of all basis, which is visible in figure 9 and 10, where it seems like $\hat{\beta}$ is not changing in the amount of noise and shows exaggerated behaviour at the boundaries. This weakness is especially pronounced in the setup $f_2,Y_1$; for the high noise setup, $f_2,Y_2$ this weakness is still visible, but less pronounced. For $f_1$, the simulation selects 5(3) and for $f_2$ 5(5) monomial basis functions for the high(low) signal-to-noise ratios. 
	\vspace{-0.2cm}

	\paragraph{B-spline Basis}
	Simulations with B-spline basis functions were possible from 4 to 18, since from 18 onwards the simulations were running into collinearity problems.
	The smooth function $f_1$ requires only 5(4) B-spline basis functions, for the high(low) signal-to-noise ratio to obtain the best fit for the B-spline basis, which performs the worst for $f_1$. An explanation might be its exaggerated behaviour at the boundaries and the exaggeration of the peculiarities of $f_1$ (figure 7 and 8), which is especially pronounced for the higher noise responses. For $f_2$, 11(6) B-spline basis functions are chosen for the high(low) signal-to-noise ratios. For $f_2$, the B-spline basis functions outperform the monomial basis but come second to the Fourier basis. While the basis seems to recognize the peculiarities in $f_2,Y_1$, it seems to struggle with it for the noisy responses in $f_2,Y_2$ (figure 9 and 10).
	\vspace{-0.2cm}

	\paragraph{Fourier Basis}
	For $f_1$, the simulation chooses a smaller number of Fourier basis functions, 5(3) and a higher number for $f_2$, 9(7) for the setup with the high(low) signal-to-noise ratio. With the low signal-to-noise ratio, the simulation chooses a smaller number of basis functions, which could be to account for the higher noise in the response variable. The Fourier basis functions perform the best for each setup for the basis expansion regression. Several reasons could contribute to this: First, especially $f_1$ shows similar curvature order across the domain while the curvature of $f_2$, while changing, does not display any erratic jumps. Second, both functions feature periodic behavior. Third, $f_2$ does have the same value at the start- and the end of the interval.	\\

	\subsubsection{Functional Principal Component Regression}
		The model which is used for the FPCR is described in \hyperref[fpc_exp_transf]{Estimation using Functional Principal Components}. Additionally to the selection of the truncation parameter $K$, the choice of the number of FPC's adds to the complexity of the model since the approximated eigenvalues and -functions from the decomposition are influenced by the choice of $K$ that was used to approximate the function. But since the FPCR is ultimately estimated with a linear model with the FPC's as regressors, the degrees of freedom in the estimation of the ultimate model are not affected by $K$, but only by the number of FPC's $n_{FPC} \in \{2, 3, 4 \}$. As will be described, it seems that neither a convex behavior of the MSPE, nor a clear relationship can be observed between the increasing number of basis function and the number of FPC's.

	\paragraph{Monomial Basis}
	For $f_1$, the cross-validated MSE is decreasing in the increasing number of FPC's and chosen basis functions for both the setup with the high, and the low signal-to-noise ratio (4,5,6 basis functions for $n_{FPC} \in \{2, 3, 4 \}$). In the setup using $f_2$ we also observe a decreasing MSE in the number of FPC, but a different behavior of the chosen basis functions: While the model selects 10 basis functions for $f_2,Y_1$ and $f_2,Y_2$ using 2 FPC, it only selects 6 for the same two specifications using 3 FPC and 8 for $f_2,Y_1$ and $f_2,Y_2$ with 4 FPC. This might be surprising, since one might assume that the number of basis functions is increasing in $n_{FPC}$ since more basis functions led to less smoothing and therefore more variability in the expanded curves that might be exploited by the FPC's. The monomial basis shows the weakest performance for all three basis functions in each setting.
	\vspace{-0.2cm}

@@

	\paragraph{Fourier Basis}
	We observe similar behaviour as for the FPCR using B-splines. In $f_1,Y_1$, the MSE is decreasing in $n_{FPC}$ while the results in $f_1,Y_2$ might indicate overfitting when using four FPC's. Since it is not possible to calculate the cross-validated MSE for three basis functions for $n_{FPC}$ = 4, the increasing number of basis functions as seen in the FPCR with B-splines can not be observed here for sure. $f_2,Y_1$ displays the greatest relative decline of MSE in the increasing number of FPC's, but in absolute values, therefore acknowledging the higher noise responses for $f_2,Y2$, the decline if MSE is similar to the decline observed for $f_2,Y_2$. Both configurations of $f_2$ are using 5,15,7 basis functions for $n_{FPC} \in \{2, 3, 4 \}$: This behaviour in increasing $n_{FPC}$ opposes the previously reported behaviour of the monomial basis and the increasing nature of the B-splines basis functions.
	\vspace{-0.2cm}

@@

	\section{Application}\label{Application}
		The application uses the methods and insights from the previous sections to predict the octane ratings of the introduced gasoline dataset.
		Although the simulation study granted valuable insights in the different methods in four different settings, it is not enough to determine the method and choice of basis of the application, since there is too much uncertainty involved. To point out some sources of uncertainty: First, opposing to the simulation setup, we do not know the true coefficient function. Visual inspection fuels the hypothesis that the estimated coefficient functions for $NIR$ look closer to $f_2$ than to $f_1$ but the insights from two functions are not enough to draw any conclusion. Second, we have no information about the signal-to-noise ratios of the measured octane numbers. Third, we made assumptions about the distribution of $\xi$ that are not relatable to this real-life application where we not know this distribution. 
		Therefore, we will run all specification again for the gasoline dataset and use the results of the simulation study towards the reading of the results. The application is designed similar to the simulation study: The 60 spectra of the gasoline dataset will be used to regress them on the reported octane numbers and evaluate the results as prediction MSE using 12 fold cross-validation with 5 elements per fold. In total, we conducted 1000 repetitions for each setting.
